Solving an Exam Scheduling Problem Using a Genetic Algorithm

### 2009-12-02 Paper/paper.tex
Paper update

\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{url}
\usepackage[table]{xcolor}
\usepackage{graphicx}
\usepackage{epstopdf}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDFMARK for TeX and GhostScript
% Uncomment and complete the following for metadata if
% your paper is typeset using TeX and GhostScript (e.g
% if you use .ps or .eps files in your paper):
% \special{! /pdfmark where
% {pop} {userdict /pdfmark /cleartomark load put} ifelse
% [ /Author (John Doe, Jane Doe)
% /Title (Paper Title)
% /Keywords (AAAI, artificial intelligence)
% /DOCINFO pdfmark}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDFINFO for PDFTeX
% Uncomment and complete the following for metadata if
% your paper is typeset using PDFTeX
\pdfinfo{
  /Title (Solving an Exam Scheduling Problem Using a Genetic Algorithm)
  /Subject (Input The Proceedings Title Here)
  /Author (Dave, Kordalewski;
           Caigu, Liu;
           Kevin, Salvesen;)
} 
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Uncomment only if you need to use section numbers
% and change the 0 to a 1 or 2
% \setcounter{secnumdepth}{0}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

    The details of the function we used are essentially arbitrary.

@@

  \subsection{Options for the Genetic Algorithm}
    \subsubsection{Scheduling Problem Instances}
      We generated 2 sets of 3 scheduling problem instances to test the algorithms above with. 
      The package \texttt{ScheduleInstanceGenerator} in our code repository manages the generation of 
      random problem instances with many free parameters to specify characteristics of the 
      output instance.

@@

      The following table characterized the problem instances that we used:

      \definecolor{light-gray}{gray}{0.87}
      \rowcolors{1}{light-gray}{white}
      \begin{center}
      \begin{small}
      \begin{tabular}{*{7}{c}}
      \hline 
      	& S 	& M 	& L 	& $S_{Cor}$ & $M_{Cor}$ & $L_{Cor}$ \\ 
      \hline \hline
       d 		& 7 	& 10	& 7 	& 7 	& 10 	& 7 	\\ 
       t 		& 5		& 8 	& 8 	& 5 	& 8 	& 8 	\\ 
       $|C|$	& 20 	& 200 	& 603 	& 20 	& 200 	& 600 	\\ 
       $|R|$	& 4 	& 10 	& 43 	& 4 	& 10 	& 43	\\ 
       $|S|$	& 50 	& 300 	& 21945 & 50 	& 300 	& 21945	\\ 
       $|L_s|$	& 1-5 	& 1-6 	& 1-5 	& 2-6 	& 2-6 	& 2-6 	\\ 
       \begin{tiny}$\#$Maj\end{tiny}	& N/A 	& N/A 	& N/A 	& 4 	& 10 	& 20 	\\ 
       \textsc{sss}	& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
      \hline 
	  \end{tabular} 
	  \end{small}
	  \end{center}

	  The large data set was created with parameters inspired from the University of 
	  Toronto's Faculty of Arts and Science.
	  The full input files for these instances are available in our repository in the 
	  directory \emph{Instance Input Files}.

    \subsubsection{Selecting Parameters for the GA}
      In order to select parameters for the genetic algorithm (that is, the population, 
      number of generations to simulate, proportions of copies, randoms, mutations, and 
      crossovers for the next generation) we performed tests using the small uncorrelated 
      problem instance.

@@

        In the graph below, each data point represents the average of 5 trials. To make 
        the scale clearer, the colour on the surface corresponds to the fitness, where 
        the highest fitnesses are red. The number of copies is fixed at 1, and the number 
        of mutations (m) and crossovers (c) are shown on the horizontal axes. The total 
        size of the population is fixed, so the number of random individuals used at each 
        point is 60?-(m+c). This corresponds to the distance from the diagonal.

        [GRAPH]

@@

      \paragraph{Size Parameters}
        Blah blah blaj. We investigated the population size and

        [Insert fancy 3D graph here = WIN]

  \subsection{Comparison Between the Different Searches}
    [\ldots]
    \begin{figure}[h!]
  	  \centering
  	  \setlength\fboxsep{0pt}
	  \setlength\fboxrule{0.5pt}
	  \fbox{\includegraphics[width=8.35cm]{HUGE-TYPE2_FullAverageStats}}
      \caption{Search comparison on a large data set}
	\end{figure}

\section{Evaluation}
  \subsection{Success}
    [\ldots]
    Seeing as our fitness function gives a very high penalty for schedules that have direct
    conflicts (either a student with two exams at the same moment, or two different exams
    in a same room at the same moment), if such schedules exists given the input data, 
    our algorithms will tend to very quickly reject them.

\section{Conclusion}
  [\ldots]
  We come up with a conclusion that in this particular problem, genetic algorithm
  is better than random search, and it is better than hill-climbing if the input 
  is a large data set. But it may not be the best solution for this problem, since
  the Mutate Search performs way better.

>>>>

\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{url}
\usepackage[table]{xcolor}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{float}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDFMARK for TeX and GhostScript
% Uncomment and complete the following for metadata if
% your paper is typeset using TeX and GhostScript (e.g
% if you use .ps or .eps files in your paper):
% \special{! /pdfmark where
% {pop} {userdict /pdfmark /cleartomark load put} ifelse
% [ /Author (John Doe, Jane Doe)
% /Title (Paper Title)
% /Keywords (AAAI, artificial intelligence)
% /DOCINFO pdfmark}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDFINFO for PDFTeX
% Uncomment and complete the following for metadata if
% your paper is typeset using PDFTeX
\pdfinfo{
  /Title (Solving an Exam Scheduling Problem Using a Genetic Algorithm)
  /Subject (Input The Proceedings Title Here)
  /Author (Dave, Kordalewski;
           Caigu, Liu;
           Kevin, Salvesen;)
} 
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Uncomment only if you need to use section numbers
% and change the 0 to a 1 or 2
% \setcounter{secnumdepth}{0}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

    We chose the weighting constants $w_s=2$ and $w_r=1$, which suggest that we are 
    more concerned with the students' schedules than the rooms'.

    The details of the function we used are essentially arbitrary, and we expect 
    similar results would be attained for any other penalty-based fitness function 
    modelling the preferences of students and requirements of rooms.

@@

    \subsection{Scheduling Problem Instances}
      We generated 2 sets of 3 scheduling problem instances to test the algorithms above with. 
      The package \texttt{Schedule Instance Generator} in our code repository manages the generation of 
      random problem instances with many free parameters to specify characteristics of the 
      output instance.

@@

      The following table characterized the problem instances that we used 
      (with \textsc{SSS} being an approximation of the search space size):      
      \definecolor{light-gray}{gray}{0.87}
      \rowcolors{1}{light-gray}{white}
      \begin{center}
      \begin{small}
      \begin{tabular}{*{7}{c}}
      \hline 
      	& S 	& M 	& L 	& $S_{Cor}$ & $M_{Cor}$ & $L_{Cor}$ \\ 
      \hline \hline
       d 		& 7 	& 10	& 7 	& 7 	& 10 	& 7 	\\ 
       t 		& 5		& 8 	& 8 	& 5 	& 8 	& 8 	\\ 
       $|C|$	& 20 	& 200 	& 603 	& 20 	& 200 	& 600 	\\ 
       $|R|$	& 4 	& 10 	& 43 	& 4 	& 10 	& 43	\\ 
       $|S|$	& 50 	& 300 	& 21945 & 50 	& 300 	& 21945	\\ 
       $|L_s|$	& 1-5 	& 1-6 	& 1-5 	& 2-6 	& 2-6 	& 2-6 	\\ 
       \begin{tiny}$\#$Maj\end{tiny}	& N/A 	& N/A 	& N/A 	& 4 	& 10 	& 20 	\\ 
       \textsc{SSS}	& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
      \hline 
	  \end{tabular} 
	  \end{small}
	  \end{center}

	  The large data set was created with parameters inspired from the University of 
	  Toronto's Faculty of Arts and Science.

\section{Evaluation}
   \subsection{Selecting Parameters for the GA}
      In order to select parameters for the genetic algorithm (that is, the population, 
      number of generations to simulate, proportions of copies, randoms, mutations, and 
      crossovers for the next generation) we performed tests using the small uncorrelated 
      problem instance.

@@

        In the graph below, each data point represents the average of 5 trials. To make 
        the scale clearer, the colour on the surface corresponds to the fitness, where 
        the highest fitnesses are red. The number of copies is fixed at 1, and the proportions 
        of mutations and crossovers are shown on the horizontal axes.

        \begin{figure}[H]
  		  \centering
  		  \includegraphics[width=8cm]{GAParametersData}
  	      \caption{Effectiveness of the GA's proportion parameters}
  	      \label{graph:GAParam1}
		\end{figure}

@@

      \paragraph{Size Parameters}
        We also performed tests to gauge the optimal population size and number of generations 
        to calculate. We determined that very small values for either the population or number 
        of generations result in very poor generated solutions. A population of around 50 is 
        sufficient for good diversity to be retained in the population, and having a number of 
        generations relatively large compared to the population size seemed to be effective.

        We gathered data from the small and medium problem instances, but the precise qualities 
        provided by particular parameters will not necessarily hold in larger instances. The 
        optimal population size and number of generations (for a fixed number of fitness 
        evaluations) is likely to vary for larger problem instances, but it seemed clear that 
        the number of generations should be set somewhat higher than the population.

        %\begin{figure}[H]
  		%  \centering
  		%  \includegraphics[width=8cm]{MEDIUM-GA_Size_Data}
  	    %  \caption{Effectiveness of the GA's population parameters}
  	    %  \label{graph:GAParam2}
		%\end{figure}

  \subsection{Comparison Between the Different Searches}
    We tested the 4 search methods on all 6 of our peoblem instances. The results for 
    only two of the problems are shown, she small and large problems, since the medium 
    problem exhibited characteristics very similar to the large problem.

    Since all methods produced nearly indistinguishable results for the uncorrelated 
    versions. This suggests that it may be possible to use the expected internal 
    correlations more effectively that the methods we have proposed do.

    On the graphs for the small instance, the results shown are averages over many trial 
    runs. Individually, however, the different methods all produced similar results, in 
    particular the ranking of the algorithms was very consistent for any particular problem 
    instance.

    \begin{figure}[H]
  	  \centering
  	  \includegraphics[width=8cm]{SMALL-TYPE2_FullAverageStats}
      \caption{Comparison of the different searches on a small data set}
	\end{figure}

	\begin{figure}[H]
  	  \centering
  	  \includegraphics[width=8cm]{HUGE-TYPE2_FullAverageStats}
      \caption{Comparison of the different searches on a large data set}
	\end{figure}

	The mutate search algorithm performed remarkably well in all tests, finding a 
	schedule at least somewhat better than every other method for every size problem 
	instance.

\section{Conclusion}
  [\ldots]
